Query Optimization for Dynamic Imputation

### 2017-05-15 paper/rebuttal.tex

\setlist[itemize]{leftmargin=*,itemsep=-0.1em}

\begin{document}
\section{Introduction}
The authors would like to thank the reviewers for their helpful and insightful comments.
We have taken steps to address each comment and we believe that the paper is significantly improved
as a result. In this document, we give a point-by-point response to the comments from each reviewer.

In the rebuttal, the reviewer's original comments are in black and our responses are in \resp{blue}.
In our draft, new or significantly changed text is highlighted in {\color{blue}blue}.

\section{Overall changes}
\begin{itemize}[leftmargin=*]
\item \resp{
    Several reviewers commented on redundant figures, excessive whitespace, and
    formatting, so we've compressed many of the figures and removed whitespace to make room for additional material.
}

\item \resp{
    We have improved our system since the paper was submitted by adding imputation for categorical variables.
    This has a minor impact on the experimental results.
    It improves the ability of the decision tree model to perform imputation by making these variables available to the model.
  }

@@

\item \resp{We extended experiments to include the new operators and provide more details on results for $\alpha$ (\Cref{sec:results}). We also added the Pareto frontiers for various plans (\Cref{fig:pareto-frontiers} ) and
explored planning times for more complex queries.}

\item \resp{We included a new section that provides a roadmap for implementing \ProjectName{} on top of an existing database. (\Cref{sec:implementation})}
\end{itemize}

  \section{Reviewer 1}
  \begin{itemize}[leftmargin=*]
  \item W2. The paper completely obscures the ML technique being used: (1) It would be nice to have some discussion of the complete API for implementing an ML technique, as well as how one might implement a cost-estimator for one or two examples. (2) It would be helpful to see the impact of a few different approaches (e.g., Tree-based, Interpolation, Regression, FD-based) on performance and accuracy. (3) Given that the ML time estimator is being treated as a black box, it would be useful to see how sensitive the plan selector is to inaccurate estimates.

  \resp{
    We extended our evaluation to include two commonly used imputation methods: mean value imputation and hot-deck imputation.
    These methods are in addition to the original imputation method which is based on decision trees.
    The decision tree model incorporates correlations between attributes, so should be able to take advantage of any functional dependencies to make better predictions.
    Interpolation based imputation is useful for data where the ordering of tuples is significant, like time series, but is not well suited for survey data where the ordering of tuples is arbitrary.

@@

      We can trivially extend \ProjectName{} to warn users when the query chosen for execution has a high $\textsc{Penalty}$ estimate, along with the number of tuples that have
      been dropped (after execution), so that situations such as those observed in Q8 can be identified by the user. We've added this detail to the paper (\Cref{sec:results}).
    }

  \item Although it doesn't consider cost or the drop operator, a similar technique appears in Yang et. al. "Lenses: An On-Demand Approach to ETL." A comparison would be nice.

   \resp{
     Yang's work is definitely related, and we've added it to our related work section.
   Their system helps an analyst prioritize cleaning tasks for a particular query.
   It uses pre-trained regression models as one of a number of repair strategies for missing data rather than creating a query plan which incorporates the training and running of such models. (\Cref{subsec:missing-value})
   }

 \item S1: ``Furthermore, the results obtained with this approach are similar in quality to those obtained with the traditional imputation approach between 0 and 24 percent" 24 percent error is pretty bad.

@@

      Avoiding this case requires us to estimate the skew in these intermediate tables, which is an area that traditional query optimizers struggle with as well.
      However, if we had improved intermediate estimates, we could incorporate them to improve our system.
      }
    %\todo{Need a better explanation here.}

  \item S3.4.4: It appears that Time() assumes that plans are IO-bound. This is a perfectly reasonable assumption, but should be stated explicitly, as datasets that fit in memory will require a different cost model.

    \resp{Agreed. Our database stores tables in files, so our plans are I/O bound, but a database which uses an in-memory representation will need a different Time() function. We have clarified this in the paper (\Cref{sec:query-runtime}).}

  \item S3.4.4: ``Drop is a special case of a sequential scan, and its time complexity can be expressed as a function of the number of heap pages read and the I/O cost per page." seems to suggest that drop is implemented as a blocking operator. Why not just take the same approach as selection and treat it as having negligible cost?

@@

  \item S4.4.1: Imputation over the raw data is being used as ground truth. Is there any reason to suspect that this is the correct thing to do? Perhaps joining multiple tables together exposes new attributes that makes imputation *more* reliable? At the very least, it would be useful to understand why there is so much error in $\alpha=0$ on Q2 and Q8.

    \resp{For our purposes, imputation over the raw data is a reasonable ground truth. This
      is because the analyst, in practice, should be considering the tradeoffs on quality
      and performance of using ImputeDB versus a production imputation system on the raw
      data. Unfortunately, they would never have access to real ground truth in such a
      scenario, and moreover, finding ground truth for use even in an empirical evaluation
      of imputation methods can be challenging because artificially dirtying the data
      often does not capture the complex reasons for missingness in real world datasets.
      We updated the discussion in \Cref{sec:results} to emphasize this point.}

  \item There are a number of opportunities for trimming in the figures, tables, and algorithms: Algorithms 1-3 could be compacted a bunch, Table 1 is full of whitespace (e.g., make it a 3-column {figure*}), Tables 1-2 could also be summarized as CDFs, Figure 6 is full of whitespace, Table 3 could be put queries on a single row (e.g., Q1 would easily fit on 1 line), and the references have a lot of redundant information (URLs, for example).

    \resp{We have compacted the algorithms, tables, and references as possible. Note that Tables 1-2 show the percentage of missing values for \emph{each attribute}, not as an overall percentage, so cannot be summarized as CDFs. We think that the URLs add value to the references, particularly for data sets.}
  \end{itemize}

@@

  \item D1. Section 3, which is the meat of the paper, is very overloaded, without getting in depth in various parts of the system. It definitely should be split in more sections, and more details should be given. In particular:
    -- Give some more information about cardinality estimation. Also, Algorithm 3 is too far from the text, and needs some explanation too.
    -- Some more details are also required in the Imputation quality section. Why are you choosing these specific formulas for Loss, especially for the impute operator. Some examples would help.
    -- Algorithms 1 and 2 could probably be presented in an easier to grasp way.

    \resp{
    We've broken up this section and consolidated algorithm notation. The cardinality estimation section provides more detail about the algorithm. We also
    added intuition for how the heuristic Penalty (a renaming of Loss, for clarity) and Time should be designed. We walk through these definitions for three different
    impute operators.
    }

  \item D2. It would be nice to show how exactly your approach can be implemented in an existing optimizer. This comes up in various places in the paper:
    -- When defining the search space, why not first define the transformation operators that inject the imputation operators. Then any optimizer can pick these rules up and add them in its plan enumeration. Of course, the used approach that follows System-R's plan enumeration is useful, but why lose generality when presenting the search space. Moreover, showing the transformation rules can help you show the size of the search space that will better justify the restrictions you apply.
    -- Why use simplified cost formulas for the traditional relational operators and not existing systems' formulas?
    -- It would be great (but not necessary of course) to add your techniques in an existing DBMS, such as Postgres to see the impact on a real system in your experiments.

   \resp{
   We've added a section (\Cref{sec:implementation}) that describes in detail the components that need to be added to an existing database system to implement our technique. 
   The details come directly from our implementation of \ProjectName{} on top of SimpleDB, a popular teaching database used at MIT, University of Washington, and Northwestern University, 
   among others. These extensions could further exploit any existing sophistication in the underlying system for tasks such as execution cost estimates.

### 2017-05-15 paper/conclusion.tex
paper: squeeze

We discuss the long history of dealing with missing data both in the statistical learning and database communities.
Similar to existing work, we consider the impact of \nullv{} values in databases and develop a simple set of invariants to 
successfully plan around them. In contrast to existing statistical learning work, our emphasis is not on the specific algorithm
used to impute but rather on the placement of imputation steps in the execution of a query. In contrast to existing database work,
we incorporate imputation into a cost-based optimizer and hide any details
regarding missing values inside the system, allowing users to use traditional SQL as if the database were complete.
While prior work has incorporated operations such as prediction into their databases, this has been domain-specific
in some cases, and in others, operations have not been integrated into the planner. The novelty
of our contribution lies in the formalization of missing value imputation for query planning, which results in performance
gains over traditional approaches. This approach acknowledges that different users performing different queries on the same
dataset will likely have varying imputation needs and that execution plans should appropriately reflect that variation.

>>>>

We discuss the long history of dealing with missing data both in the statistics and database communities.
In contrast to existing work in statistics, our emphasis is not on the specific algorithm
used to impute but rather on the placement of imputation steps in the execution of a query. In contrast to existing database work,
we incorporate imputation into a cost-based optimizer and hide any details
regarding missing values inside the system, allowing users to use traditional SQL as if the database were complete.
While prior work has incorporated operations such as prediction into their databases, this has been domain-specific
in some cases, and in others, operations have not been integrated into the planner. The novelty
of our contribution lies in the formalization of missing value imputation for query planning, resulting in performance
gains over traditional approaches. This approach acknowledges that different users performing different queries on the same
dataset will likely have varying imputation needs and that execution plans should appropriately reflect that variation.
